---
title: "Panel Analysis of the impact of WF Smoke on Solar Gen: West US 2013-2020"
subtitle: "Parking Lot"
output: 
  html_document: default
  md_document:
    variant: markdown_github
---
~~~~~~~ Parking Lot ~~~~~~

```{r "Pulling PM2.5 monitor characteristics from EPA", eval=FALSE}

#Define key parameters
EPAkey <- "blueswift59"
parameter <- "88101"  #Returns local PM 2.5 measures
bdate <- "20200101"
edate <- "20201231"
frequency <- "monthly"
state <- "06"
county <- c("067","013","005","017","009","061","077","095","113","101")
n_iter <- 10
  
# Initializes the progress bar
pb <- txtProgressBar(min = 0,
                     max = n_iter,
                     style = 3,
                     width = 50,
                     char = "=")   

j=0
for(i in 1:10) {
  
  url_epa <- 
    paste("https://aqs.epa.gov/data/api/monitors/byCounty?email=brian@thecgo.org",
          "&key=",EPAkey,
          "&param=",parameter,
          "&bdate=",bdate,
          "&edate=",edate,
          "&state=",state,
          "&county=",county[i],
          sep = "")

  res_epa <- GET(url_epa)

  pm25AZ <- fromJSON(rawToChar(res_epa$content))
  
  pm25AZ_df <- data.frame(pm25AZ$Data)
  
    if (j == 0){
      pm25AZ_final <- pm25AZ_df
   }
    else{
      pm25AZ_final <- rbind(pm25AZ_final, pm25AZ_df)
    }
  
  setTxtProgressBar(pb, i)
  
  j = j+1

}

close(pb)

plant_char <- subset(plant_char, 
                       select = c(
                         "sector",
                         "entityid",
                         "entityName",
                         "plantid",
                         "balancing_authority_code",
                         "status",
                         "county",
                         "latitude",
                         "longitude",
                         "nameplate.capacity.mw",
                         "operating.year.month"
    
))


```

```{r  "Define functions to pull air monitor data using EPA's API"}

# Define a function to call the PM2.5 hourly data using EPA's API ----------

pm25.sample.api.call <-
  function(bdate,
           edate,
           state,
           EPAkey = "blueswift59",
           parameter = "88101",
           frequency = "monthly",
           duration = "1",
           ...) {
    url_epa <-
      paste(
        "https://aqs.epa.gov/data/api/sampleData/byState?email=brian@thecgo.org",
        "&key=",
        EPAkey,
        "&param=",
        parameter,
        "&bdate=",
        bdate,
        "&edate=",
        edate,
        "&state=",
        state,
        "&duration=",
        duration,
        sep = ""
      )
    
    res_epa <- GET(url_epa)
    
    pm <- fromJSON(rawToChar(res_epa$content))
    
    pm_final <- data.frame(pm$Data)
    
    return(pm_final)
  }

pm25.daily.api.call <-
  function(date,
           state,
           EPAkey = "blueswift59",
           parameter = "88101",
           ...) {
    url_daily_epa <-
      paste(
        "https://aqs.epa.gov/data/api/dailyData/byState?email=brian@thecgo.org",
        "&key=",
        EPAkey,
        "&param=",
        parameter,
        date,
        "&state=",
        state,
        sep = ""
      )
    
    res_daily_epa <- GET(url_daily_epa)
    
    pm_daily <- fparse(res_daily_epa$content)
    
    pm_daily_final <- subset(
      pm_daily$Data,
      select = c(
        "state_code",
        "county_code",
        "site_number",
        "poc",
        "latitude",
        "longitude",
        "sample_duration_code",
        "sample_duration",
        "date_local",
        "event_type",
        "observation_count",
        "observation_percent",
        "validity_indicator",
        "arithmetic_mean",
        "aqi",
        "method_code",
        "method",
        "local_site_name",
        "state",
        "county",
        "city"
      )
    )
      
    return(pm_daily_final)
  }

# Define a function to return monthly PM2.5 averages ----------

monthly.avg.pm <- function(df, ...) {
  df$day <- substr(df$date_local, 9, 10)
  df$month <- substr(df$date_local, 6, 7)
  df$year <- substr(df$date_local, 1, 4)
  
  df <- df %>% drop_na(sample_measurement)
  
  result <- df %>% group_by(state, county, year, month) %>%
    summarize(pm_measure = mean(sample_measurement))
  
  return(result)
  
}

monthly.avg.daily.pm <- function(df, ...) {
  df$day <- substr(df$date_local, 9, 10)
  df$month <- substr(df$date_local, 6, 7)
  df$year <- substr(df$date_local, 1, 4)
  
  result <- df %>% group_by(state, county, year, month) %>%
    summarize(pm_measure = mean(arithmetic_mean))
  
  return(result)
  
}

# Define a function to return daily PM2.5 averages ----------

daily.avg.pm <- function(df, ...) {
  df$day <- substr(df$date_local, 9, 10)
  df$month <- substr(df$date_local, 6, 7)
  df$year <- substr(df$date_local, 1, 4)
  
  df <- df %>% drop_na(sample_measurement)
  
  result <-
    df %>% group_by(state_code, county_code, state, county, year, month, day) %>%
    summarize(pm_measure = mean(sample_measurement))
  
  return(result)
  
} 
```

```{r EIA API v1 example, eval=FALSE}

#Call multiple series using a for loop
j = 0
for(i in plantID.AZ$Plant.Id) {
  
  url <- 
    paste("https://api.eia.gov/series/?api_key=",EIAkey,
          "&series_id=ELEC.PLANT.GEN.",i,"-SUN-PV.M",sep = ""
               )
  
  res <- GET(url)
  
  json_data.AZ <- fromJSON(rawToChar(res$content), flatten = TRUE
                        )
  APIdata.AZ <- data.frame(json_data.AZ$series$data)
  
  #Create new columns to organize the data we want
  APIdata.AZ$Year <- substr(APIdata.AZ$X1,1,4) 
  APIdata.AZ$Month <- substr(APIdata.AZ$X1,5,6)
  APIdata.AZ$Day <- 1
  APIdata.AZ$Date <- as.Date(paste(APIdata.AZ$Year, APIdata.AZ$Month, APIdata.AZ$Day, sep="-"))
  colnames(APIdata.AZ)[2] <- word(json_data.AZ$series$name, 1, sep = fixed(":"))
  #APIdata.AZ$Lat <- json_data.AZ$series$lat
  #APIdata.AZ$Lon <- json_data.AZ$series$lon
  
  APIdata.AZ <- APIdata.AZ[-c(1,5)]
  
  if (j == 0){
    APIdata.AZ.final <-APIdata.AZ
  }
  else{
    APIdata.AZ.final <- merge(APIdata.AZ.final, APIdata.AZ, by="Date")
  }
  
  j = j+1
}

#APIdata.AZ.final <- subset(APIdata.AZ.final, Date>= startdate & Date <= enddate)

write.csv(APIdata.AZ.final,"temp/APIdata.AZ.csv")
write.csv(json_data.AZ,"temp/json_data.AZ.csv")
```

Import 860 data from 2___Plant_Yyyyy and 3_3_Solar_Yyyyy (only operable plants)

```{r, eval=FALSE}

plant2020 <- read.csv("data/raw/form_860/EIA860_2020_plant.csv")

solar2020 <- read.csv("data/raw/form_860/EIA860_2020_solar.csv")

```

Remove unnecessary columns from plant2020

```{r, eval=FALSE}

plant2020 <- subset(plant2020, select = c(Plant.Code,
                                          Plant.Name,
                                          Street.Address,
                                          City,
                                          State,
                                          Zip,
                                          County,
                                          Latitude,
                                          Longitude
))

```
Combine form 860 data using Plant Id

```{r, eval=FALSE}

plantchar2020 <- inner_join(x = solar2020, y = plant2020, by = "Plant.Code")

```

Remove unwanted info by creating a subset of imported data

Check if new joined dataset is matched correctly. If utility IDs and counties match,then duplicates are removed and a simpler dataset is created.

```{r, eval=FALSE}

if(all(plantchar2020$Utility.ID.x == plantchar2020$Utility.ID.y)
   &(all(plantchar2020$County.x == plantchar2020$County.y))
    ){simple_plantchar2020 <- plantchar2020[!duplicated(as.list(plantchar2020))]
    }else print("MISMATCH")

```

Remove unwanted columns

```{r, eval=FALSE}

plantchar2020 <- subset(plantchar2020, 
                        select = -c(Ash.Impoundment., 
                                    Ash.Impoundment.Lined.,
                                    Ash.Impoundment.Status, 
                                    Natural.Gas.LDC.Name,
                                    Natural.Gas.Pipeline.Name.1, 
                                    Natural.Gas.Pipeline.Name.2,
                                    Natural.Gas.Pipeline.Name.3, 
                                    Pipeline.Notes,
                                    Natural.Gas.Storage,
                                    Liquefied.Natural.Gas.Storage
))

```

Import form 923 generation data for years 2013-2020:

Create a vector for each file name
(NOTE: THE 2013 DATA HAS SLIGHTLY DIFFERENT COLUMN HEADINGS, I CHANGED THEM
MANUALLY IN EXCEL TO MATCH 2014-2020)

```{r, eval=FALSE}

form923_paths <- fs::dir_ls("data/raw/form_923")
form923_paths

```

Write a loop that will read in all files

```{r, eval=FALSE}

list_form923 <- list()

for (i in seq_along(form923_paths)) {
    list_form923[[i]] <- read.csv(
      file = form923_paths[[i]], fileEncoding = 'UTF-8-BOM'
    )
}

list_form923 <- set_names(list_form923, form923_paths)

```

Eliminate all non-utility and non-solar generation from each

```{r, eval=FALSE}

pvlist_form923 <- lapply(list_form923, function(x)
                         filter(x, NAICS.Code == "22",
                                Reported.Prime.Mover == "PV"))

```

Get rid of unnecessary columns of data

```{r, eval=FALSE}

simplepvlist_form923 <- lapply(pvlist_form923, function(x)
                              subset(x, select = c(Plant.Id, 
                                                   Plant.Name, 
                                                   Plant.State, 
                                                   Census.Region, 
                                                   NERC.Region, 
                                                   Netgen.January, 
                                                   Netgen.February,	
                                                   Netgen.March, 
                                                   Netgen.April,	
                                                   Netgen.May, 
                                                   Netgen.June,	
                                                   Netgen.July, 
                                                   Netgen.August,	
                                                   Netgen.September,	
                                                   Netgen.October,	
                                                   Netgen.November,	
                                                   Netgen.December,	
                                                   YEAR)))

```

Combine each annual form 923  data set to form one complete table 

```{r, eval=FALSE}

monthlygen <- simplepvlist_form923 %>% reduce(full_join, by="Plant.Id")
write.csv(monthlygen, "temp/monthlygen.csv")

```

Combine form 923 data with form 860 data by matching plant code with plant id

```{r, eval=FALSE}

monthlygen2 <- left_join(x = monthlygen, y = plant2020, by = c("Plant.Id" = "Plant.Code"))
write.csv(monthlygen2, "temp/monthlygen2.csv")

```

Attempt to convert from wide to long

```{r, eval=FALSE}

setDT(monthlygen2)
longmonthlygen <- melt(monthlygen2, id.vars=c("County"))
write.csv(longmonthlygen, "temp/longmonthlygen.csv")

```

Remove duplicates created by multiple generator ids

```{r, eval=FALSE}

spv2020data <- distinct(pv2020data, Plant.Id, .keep_all = TRUE)

```

Remove unnecessary data columns

```{r, eval=FALSE}

cpv2020data <- subset(spv2020data, select = c(Plant.Id, Plant.Name,
                                             Census.Region, NERC.Region.x,
                                             Respondent.Frequency, Netgen.January,	
                                             Netgen.February,	Netgen.March,	Netgen.April,	
                                             Netgen.May,	Netgen.June,	Netgen.July,	
                                             Netgen.August,	Netgen.September,	Netgen.October,	
                                             Netgen.November,	Netgen.December,	
                                             YEAR,	State.x,
                                             County.x,	Generator.ID,	Status,
                                             Nameplate.Capacity..MW.,	
                                             Operating.Year,	Street.Address,	
                                             City,	Zip,	Latitude,	Longitude))

```

Set cpv2020data as a data table

```{r, eval=FALSE}

setDT(cpv2020data)

```

Convert data from wide to long form

```{r, eval=FALSE}

Wide2020 <- melt(cpv2020data,
                 measure.vars = c("Netgen.January",	"Netgen.February","Netgen.March",	
                                  "Netgen.April",	"Netgen.May",	"Netgen.June","Netgen.July",	
                                  "Netgen.August","Netgen.September","Netgen.October",	
                                  "Netgen.November",	"Netgen.December"),
                 variable.name = "Month",
                 value.name = "Generation")

```