---
title: "Panel Analysis of the impact of WF Smoke on Solar Gen: West US 2013-2020"
output: 
  html_document: default
  md_document:
    variant: markdown_github
---

R Markdown cheat sheet: https://www.rstudio.com/blog/the-r-markdown-cheat-sheet/

Use "?rmarkdown::_________" to look up formatting options for the YAML
header. For example, "?rmarkdown::html_document" will provide potential options
for html documents. More info on Rmarkdown anatomy can be found at
https://bookdown.org/yihui/rmarkdown-cookbook/rmarkdown-anatomy.html

Short guide to chunk options: https://bookdown.org/yihui/rmarkdown/r-code.html


#Cleaning and Combining the Data


Solar powerplant characteristics:
Download EIA form 860 data from https://www.eia.gov/electricity/data/eia860/

Monthly plant-level generation data
Download EIA for 923 data from https://www.eia.gov/electricity/data/eia923/

Install and load packages

```{r, include=FALSE}

library(easypackages)
libraries("readxl","ggplot2","stringr","rlist","data.table","DBI","RSQLite",  
          "stargazer","Hmisc","magrittr","formattable","RPostgreSQL", 
          "tidyverse","dbplyr","pbapply","stargazer","quantreg","olsrr", 
          "faraway","binaryLogic","sf","caret","Amelia","maps", 
          "tidycensus","reshape","boot","biglm","tigris","summarytools",
          "dplyr","fs", "httr","jsonlite")

options(scipen=999)
```

##Monthly Solar Generation

Data for monthly solar generation is collected via the US Energy Information Agency's Open Data Application Programming Interface (API): https://www.eia.gov/opendata/

Plant level data can be found by clicking EIA Data Sets > Electricity > Plant level data.

In order to pull in the necessary data, however, we first need to know the Plant ID for each solar power plant we wish to analyze. We can pull those IDs from EIA's form 860: https://www.eia.gov/electricity/data/eia860/

```{r Utility-scale PV Plant IDs}

plantID.AZ <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "AZ")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.CA <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "CA")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.CO <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "CO")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.ID  <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "ID")%>%
  select(c("Plant.Id"))

plantID.ID.list <- plantID.ID$Plant.Id %>%
  as.list()

plantID.NM <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "NM")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.NV <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "NV")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.OR <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "OR")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.TX <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "TX")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.UT <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "UT")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.WA <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "WA")%>%
  select(c("Plant.Id"))%>%
  as.list()

plantID.WY <- fread(file = "data/raw/form_923/EIA923_2020_monthlygen.csv",
                 keepLeadingZeros = TRUE,
                 check.names = TRUE,
                 colClasses = "character")%>%
  filter(NAICS.Code == "22",
         Reported.Prime.Mover == "PV",
         Netgen.August != "0",
         Plant.State == "WY")%>%
  select(c("Plant.Id"))%>%
  as.list()
```

Now we can use the list of plant IDs in a for loop to create a data frame of monthly solar generation

```{r EIA API Call}

#Define variables
EIAkey <- "iPjXaaJSiyfWNALV62yg1Onb0VTuUgVHa0eVHgbx"
startdate <- "2013-01-01"
enddate <- "2020-12-31"

#Call multiple series using a for loop

j=0
for(i in plantID.ID.list) {
  
  url <- 
    paste("https://api.eia.gov/series/?api_key=",EIAkey,
          "&series_id=ELEC.PLANT.GEN.",i,"-SUN-PV.M",sep = ""
               )
  
  res <- GET(url)
  
  json_data.ID <- fromJSON(rawToChar(res$content), flatten = TRUE
                        )
  APIdata.ID <- data.frame(json_data.ID$series$data)
  
  #Create new columns to organize the data we want
  APIdata.ID$Year <- substr(APIdata.ID$X1,1,4) 
  APIdata.ID$Month <- substr(APIdata.ID$X1,5,6)
  APIdata.ID$Day <- 1
  
  APIdata.ID$Date <- as.Date(paste(APIdata.ID$Year, APIdata.ID$Month, APIdata.ID$Day, sep="-"))
  colnames(APIdata.ID)[2] <- word(json_data.ID$series$name, 1, sep = fixed(":"))
  #APIdata.ID$Lat <- json_data.ID$series$lat
  #APIdata.ID$Lon <- json_data.ID$series$lon
  
  APIdata.ID <- APIdata.ID[-c(1,6)]
  
   if (j == 0){
    APIdata.ID.final <-APIdata.ID
  }
  else{
    APIdata.ID.final <- merge(APIdata.ID.final, APIdata.ID, by="Date")
  }
  
  j = j+1
  
}

write.csv(APIdata.ID,"temp/APIdata.ID.final.csv")
write.csv(json_data.ID,"temp/json_data.ID.csv")
```

```{r}

#Call multiple series using a for loop
j = 0
for(i in plantID.AZ$Plant.Id) {
  
  url <- 
    paste("https://api.eia.gov/series/?api_key=",EIAkey,
          "&series_id=ELEC.PLANT.GEN.",i,"-SUN-PV.M",sep = ""
               )
  
  res <- GET(url)
  
  json_data.AZ <- fromJSON(rawToChar(res$content), flatten = TRUE
                        )
  APIdata.AZ <- data.frame(json_data.AZ$series$data)
  
  #Create new columns to organize the data we want
  APIdata.AZ$Year <- substr(APIdata.AZ$X1,1,4) 
  APIdata.AZ$Month <- substr(APIdata.AZ$X1,5,6)
  APIdata.AZ$Day <- 1
  APIdata.AZ$Date <- as.Date(paste(APIdata.AZ$Year, APIdata.AZ$Month, APIdata.AZ$Day, sep="-"))
  colnames(APIdata.AZ)[2] <- word(json_data.AZ$series$name, 1, sep = fixed(":"))
  #APIdata.AZ$Lat <- json_data.AZ$series$lat
  #APIdata.AZ$Lon <- json_data.AZ$series$lon
  
  APIdata.AZ <- APIdata.AZ[-c(1,5)]
  
  if (j == 0){
    APIdata.AZ.final <-APIdata.AZ
  }
  else{
    APIdata.AZ.final <- merge(APIdata.AZ.final, APIdata.AZ, by="Date")
  }
  
  j = j+1
}

#APIdata.AZ.final <- subset(APIdata.AZ.final, Date>= startdate & Date <= enddate)

write.csv(APIdata.AZ.final,"temp/APIdata.AZ.csv")
write.csv(json_data.AZ,"temp/json_data.AZ.csv")
```

Import 860 data from 2___Plant_Yyyyy and 3_3_Solar_Yyyyy (only operable plants)

```{r}

plant2020 <- read.csv("data/raw/form_860/EIA860_2020_plant.csv")

solar2020 <- read.csv("data/raw/form_860/EIA860_2020_solar.csv")

```

Remove unnecessary columns from plant2020

```{r}

plant2020 <- subset(plant2020, select = c(Plant.Code,
                                          Plant.Name,
                                          Street.Address,
                                          City,
                                          State,
                                          Zip,
                                          County,
                                          Latitude,
                                          Longitude
))

```
Combine form 860 data using Plant Id

```{r, eval=FALSE}

plantchar2020 <- inner_join(x = solar2020, y = plant2020, by = "Plant.Code")

```

Remove unwanted info by creating a subset of imported data

Check if new joined dataset is matched correctly. If utility IDs and counties match,then duplicates are removed and a simpler dataset is created.

```{r, eval=FALSE}

if(all(plantchar2020$Utility.ID.x == plantchar2020$Utility.ID.y)
   &(all(plantchar2020$County.x == plantchar2020$County.y))
    ){simple_plantchar2020 <- plantchar2020[!duplicated(as.list(plantchar2020))]
    }else print("MISMATCH")

```

Remove unwanted columns

```{r, eval=FALSE}

plantchar2020 <- subset(plantchar2020, 
                        select = -c(Ash.Impoundment., 
                                    Ash.Impoundment.Lined.,
                                    Ash.Impoundment.Status, 
                                    Natural.Gas.LDC.Name,
                                    Natural.Gas.Pipeline.Name.1, 
                                    Natural.Gas.Pipeline.Name.2,
                                    Natural.Gas.Pipeline.Name.3, 
                                    Pipeline.Notes,
                                    Natural.Gas.Storage,
                                    Liquefied.Natural.Gas.Storage
))

```

Import form 923 generation data for years 2013-2020:

Create a vector for each file name
(NOTE: THE 2013 DATA HAS SLIGHTLY DIFFERENT COLUMN HEADINGS, I CHANGED THEM
MANUALLY IN EXCEL TO MATCH 2014-2020)

```{r}

form923_paths <- fs::dir_ls("data/raw/form_923")
form923_paths

```

Write a loop that will read in all files

```{r}

list_form923 <- list()

for (i in seq_along(form923_paths)) {
    list_form923[[i]] <- read.csv(
      file = form923_paths[[i]], fileEncoding = 'UTF-8-BOM'
    )
}

list_form923 <- set_names(list_form923, form923_paths)

```

Eliminate all non-utility and non-solar generation from each

```{r}

pvlist_form923 <- lapply(list_form923, function(x)
                         filter(x, NAICS.Code == "22",
                                Reported.Prime.Mover == "PV"))

```

Get rid of unnecessary columns of data

```{r}

simplepvlist_form923 <- lapply(pvlist_form923, function(x)
                              subset(x, select = c(Plant.Id, 
                                                   Plant.Name, 
                                                   Plant.State, 
                                                   Census.Region, 
                                                   NERC.Region, 
                                                   Netgen.January, 
                                                   Netgen.February,	
                                                   Netgen.March, 
                                                   Netgen.April,	
                                                   Netgen.May, 
                                                   Netgen.June,	
                                                   Netgen.July, 
                                                   Netgen.August,	
                                                   Netgen.September,	
                                                   Netgen.October,	
                                                   Netgen.November,	
                                                   Netgen.December,	
                                                   YEAR)))

```

Combine each annual form 923  data set to form one complete table 

```{r}

monthlygen <- simplepvlist_form923 %>% reduce(full_join, by="Plant.Id")
write.csv(monthlygen, "temp/monthlygen.csv")

```

Combine form 923 data with form 860 data by matching plant code with plant id

```{r}

monthlygen2 <- left_join(x = monthlygen, y = plant2020, by = c("Plant.Id" = "Plant.Code"))
write.csv(monthlygen2, "temp/monthlygen2.csv")

```

Attempt to convert from wide to long

```{r, eval=FALSE}

setDT(monthlygen2)
longmonthlygen <- melt(monthlygen2, id.vars=c("County"))
write.csv(longmonthlygen, "temp/longmonthlygen.csv")

```

Remove duplicates created by multiple generator ids

```{r, eval=FALSE}

spv2020data <- distinct(pv2020data, Plant.Id, .keep_all = TRUE)

```

Remove unnecessary data columns

```{r, eval=FALSE}

cpv2020data <- subset(spv2020data, select = c(Plant.Id, Plant.Name,
                                             Census.Region, NERC.Region.x,
                                             Respondent.Frequency, Netgen.January,	
                                             Netgen.February,	Netgen.March,	Netgen.April,	
                                             Netgen.May,	Netgen.June,	Netgen.July,	
                                             Netgen.August,	Netgen.September,	Netgen.October,	
                                             Netgen.November,	Netgen.December,	
                                             YEAR,	State.x,
                                             County.x,	Generator.ID,	Status,
                                             Nameplate.Capacity..MW.,	
                                             Operating.Year,	Street.Address,	
                                             City,	Zip,	Latitude,	Longitude))

```

Set cpv2020data as a data table

```{r, eval=FALSE}

setDT(cpv2020data)

```

Convert data from wide to long form

```{r, eval=FALSE}

Wide2020 <- melt(cpv2020data,
                 measure.vars = c("Netgen.January",	"Netgen.February","Netgen.March",	
                                  "Netgen.April",	"Netgen.May",	"Netgen.June","Netgen.July",	
                                  "Netgen.August","Netgen.September","Netgen.October",	
                                  "Netgen.November",	"Netgen.December"),
                 variable.name = "Month",
                 value.name = "Generation")

```